I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-16 Paper/Paper_noProof.tex
data
sketch of proof

\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\vW}{\boldsymbol{W}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vb}{\boldsymbol{b}}
\newcommand{\val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}

@@

% testtesttesttest
An $\ell$-layer DNN is provided by $\ell$ weight matrices 
$\boldsymbol{W}^i \in \mathbb{R}^{d_i\times d_{i-1}}$
and $\ell$ bias vectors $\boldsymbol{b}^i \in \mathbb{R}^{d_i}$, for $i=1, \ldots, \ell$.
We call $d_i$ the number of neurons of hidden layer $i \in \{1, \ldots, \ell-1\}$,
$d_0$ the input dimension, and $d_\ell$ the output dimension.

Given an input vector $\boldsymbol{z}^0 \in \mathbb{R}^{d_0}$, 
denoting $\hat{\boldsymbol{z}}^{0}={\boldsymbol{z}}^0$, we define inductively the weight vector $\boldsymbol{z}^i$ at layer $1 \leq i \leq \ell$ with
\begin{align*}
	{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \boldsymbol{b}^i\\
	\hat{\boldsymbol{z}}^{i} &= ReLU({\boldsymbol{z}}^i).
\end{align*}

@@

\iffalse
and the $i$-th hidden layer is a vector in $\mathbb{R}^{d_i}$, 
and the output layer is a vector in $\mathbb{R}^{d'}$ or a scale. 
The weights, bias and activation functions decide propagate the from previous to the next layer. In formula, from layer $l_{i-1}$ to layer $l_{i}$, the weight 
$\boldsymbol{W}^i$ is matrix of $d_i\times d_{i-1}$, 
the bias is a vector $\boldsymbol{b}^i$ in $\mathbb{R}^{d_i}$, and the activation function 
is $\sigma$, then  if the $i-1$-th layer is $\hat{\boldsymbol{z}}^{(i-1)}$, 
then the value of $i$-th layer is computed by: 
\begin{align*}
	{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \boldsymbol{b}^i\\
	\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)).
\end{align*} The vector $\hat{\boldsymbol{z}}$ is called post-activation values, and $\boldsymbol{z}$ is called pre-activation values, and $\boldsymbol{z}^{(i)}_j$ is used to call the $j$-th neuron in the $i$-th layer. In our style, we also call neurons \emph{nodes} and use $a,b,c,d$ to denote them. We use $W_{ab}$ to denote the weight from neuron $b$ to $a$. We use $\boldsymbol{x}$ to denote the vector of input and  $f(\boldsymbol{x})$ to denote the output.
\fi

@@

It thus gives an upper bound at 
$\beta \cdot a = \beta \frac{y - \alpha}{\beta-\alpha}$, which is exactly the function used in DeepPoly.
\qed
\end{proof}

@@

The proof can be found in Section \ref{sec.proofs}. Notice that standard {\em DeepPoly} does not satisfies the following theorem (it does not necessarily refine {\em Box}), and therefore we are not sure PRIMA or $\beta$-CROWN who refine DeepPoly do, though it would be easy to fix by using $\overline{\text{DeepPoly}}$. Anyway, in practice, it is very unlikely to have a network without any compensating path, it is more of a theoretical results showin that compensation is the reason for inacuracy.

@@

Similarly, the proof of Theorem \ref{th2} can be found in Section \ref{sec.proofs}.

@@

Intuitively, our first proof shows that if there are no compensation, then there are no correlations between nodes, and one can maximize/minimize independently the different input nodes in order to optimize any desired node.

@@

	Constant $\boldsymbol{W}, \boldsymbol{b}$ \tcp*{Weights and Bias of network}

@@

			$\mathcal{D}$, $\boldsymbol{W}, \boldsymbol{b}$ $\rightarrow$ $\mathcal{B}_0$ = \{$x$: $(UB(x),LB(x))$,$\cdots$\} 
			\tcp*{Do linear transformation  from input to the first layer}

@@

				$\mathcal{B}$, $\boldsymbol{W}, \boldsymbol{b}$, $\mathcal{O}$ $\rightarrow \mathcal{M}$  \tcp*{MILP model $\mathcal{M}$ for node $x$}

@@

		$6\times500$\  &   $26\%$ & $35\%$ & $41\%$ & $44.5\%$  & $\mathbf{65\%}$ & $92\%$ \\

		$\epsilon = 0.035$ & 34s & 121s & 18.4s & 576s & 925s &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}
	\vspace{-0.8cm}
\end{table}

>>>>

\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\vW}{\boldsymbol{W}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vb}{\boldsymbol{bias}}
\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}

\newcommand{\UB}{\mathrm{UB}}
\newcommand{\LB}{\mathrm{LB}}
\newcommand{\ub}{\mathrm{ub}}
\newcommand{\lb}{\mathrm{lb}}
\newcommand{\B}{\mathrm{B}}

@@

% testtesttesttest
An $\ell$-layer DNN is provided by $\ell$ weight matrices 
$\boldsymbol{W}^i \in \mathbb{R}^{d_i\times d_{i-1}}$
and $\ell$ bias vectors $\vb^i \in \mathbb{R}^{d_i}$, for $i=1, \ldots, \ell$.
We call $d_i$ the number of neurons of hidden layer $i \in \{1, \ldots, \ell-1\}$,
$d_0$ the input dimension, and $d_\ell$ the output dimension.

Given an input vector $\boldsymbol{z}^0 \in \mathbb{R}^{d_0}$, 
denoting $\hat{\boldsymbol{z}}^{0}={\boldsymbol{z}}^0$, we define inductively the weight vector $\boldsymbol{z}^i$ at layer $1 \leq i \leq \ell$ with
\begin{align*}
	{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \vb^i\\
	\hat{\boldsymbol{z}}^{i} &= ReLU({\boldsymbol{z}}^i).
\end{align*}

@@

\iffalse
and the $i$-th hidden layer is a vector in $\mathbb{R}^{d_i}$, 
and the output layer is a vector in $\mathbb{R}^{d'}$ or a scale. 
The weights, bias and activation functions decide propagate the from previous to the next layer. In formula, from layer $l_{i-1}$ to layer $l_{i}$, the weight 
$\boldsymbol{W}^i$ is matrix of $d_i\times d_{i-1}$, 
the bias is a vector $\vb^i$ in $\mathbb{R}^{d_i}$, and the activation function 
is $\sigma$, then  if the $i-1$-th layer is $\hat{\boldsymbol{z}}^{(i-1)}$, 
then the value of $i$-th layer is computed by: 
\begin{align*}
	{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \vb^i\\
	\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)).
\end{align*} The vector $\hat{\boldsymbol{z}}$ is called post-activation values, and $\boldsymbol{z}$ is called pre-activation values, and $\boldsymbol{z}^{(i)}_j$ is used to call the $j$-th neuron in the $i$-th layer. In our style, we also call neurons \emph{nodes} and use $a,b,c,d$ to denote them. We use $W_{ab}$ to denote the weight from neuron $b$ to $a$. We use $\boldsymbol{x}$ to denote the vector of input and  $f(\boldsymbol{x})$ to denote the output.
\fi

@@

It thus gives an upper bound at 
$\beta \cdot a = \beta \frac{y - \alpha}{\beta-\alpha}$, which is exactly the function used in DeepPoly.
\hfill $\square$
\end{proof}

@@

The sketch of proof can be found in Section \ref{sec.proofs}. 
Notice that standard {\em DeepPoly} does not satisfies the following theorem (it does not necessarily refine {\em Box}), and therefore we are not sure PRIMA or $\beta$-CROWN who refine DeepPoly do, though it would be easy to fix by using $\overline{\text{DeepPoly}}$. Anyway, in practice, it is very unlikely to have a network without any compensating path, it is more of a theoretical results showin that compensation is the reason for inacuracy.

@@

The sketch of proof of Theorem \ref{th2} can be found in Section \ref{sec.proofs}.

@@

Intuitively, our first proof shows that if there are no compensation, then there are no correlations between nodes.
Consider a target neuron $z$.
In case there is no compensating path, we can assign a sign to each neuron $n$: 0
if all paths from $n$ to $z$ has weight 0, $+1$ if all paths have positive weights, and 
$-1$ if they have negative weights. Then we prove:

\begin{proposition}
	\label{prop.sign}
There exists two inputs $\vx^\star,\vx^\sharp$ such that:
\begin{enumerate} 
	\item ${\vx^\star}$ maximizes the value of all positive neurons and minimizes the value of all negative neurons.
	\item ${\vx^\sharp}$ minimizes the value of all positive neurons and maximizes the value of all negative neurons.
\end{enumerate}
\end{proposition}

Hence $\val_{\vx^\star}(z)=\beta$ and Hence $\val_{\vx^\sharp}(z)=\alpha$,
for $[\alpha, \beta]$ the bounds computed by the Box abstraction. We conclude Theorem \ref{th1} by invocking continuity.

\smallskip

Concerning the Theorem \ref{th2}, consider an output node $z$.
Let $[\alpha(z),\beta(z)]$ be the bound computed by $\MILP_{X_z}$.
We will show the existence of $\vx^\sharp,\vx^\star$
such that $\val_{\vx^\sharp}(z)=\alpha$ and $\val_{\vx^\star}(z)=\beta$,
and we will get the proof of Theorem 2 by a continuity argument.
For that, we partition the input neurons from layer $0$ into four groups:
\begin{enumerate}
	\item $A_{zero}$ is the set of input neurones from where all paths with target $z$ have weight $0$.
	\item $A_{pos}$ is the set of input neurons from where all paths with target $z$ have weight $\geq 0$.
	\item $A_{neg}$ is the set of input neurons from where all paths with target $z$ have weight $\leq 0$.
	\item $A_{open}$ is the set of remaining input neurons.
\end{enumerate}

We define $A_{pure}=A_{pos} \cup A_{neg}$.
We then partition the set of neurons in hidden layers: 
\begin{enumerate}
	\item $B_{zero}$ is the set of neurones from where all paths with target $z$ have weight $0$.
	\item $B_{open}$ is the set of neurones reachable from $A_{open}$ and such that there is at least one path to $z$ with non zero weight.
	\item $B_{pure}$ is the set of remaining neurons.
\end{enumerate}

Notice that the value of any node $n$ in $A_{pure}$ or $B_{pure}$ is entirely determined by 
nodes in $A_{pure}$ or $B_{pure}$. 
We can define the sign function for these neurons.
Therefore, applying Prop. \ref{prop.sign}, we obtain 
$\vx_{pure}^\star,\vx_{pure}^\sharp$ optimizing neurons in $A_{pure}$ or $B_{pure}$.
Now, assume that $\MILP_{X_z}$ computes bounds $[\alpha,\beta]$ for $z$.
We prove that there exists input 
$\vx^\star$ extending $\vx_{pure}^\star$
and input $\vx_{pure}^\sharp$ extending $\vx_{pure}^\sharp$
from $A_{pure}$ to $A_{pure} \cup A_{open}$ such that 
$\val_{\vx^\star}(z)=\beta$ and $\val_{\vx^\sharp}(z)=\alpha$.

The formal proofs are in appendix.

@@

	Constant $\boldsymbol{W}, \vb$ \tcp*{Weights and Bias of network}

@@

			$\mathcal{D}$, $\boldsymbol{W}, \vb$ $\rightarrow$ $\mathcal{B}_0$ = \{$x$: $(UB(x),LB(x))$,$\cdots$\} 
			\tcp*{Do linear transformation  from input to the first layer}

@@

				$\mathcal{B}$, $\boldsymbol{W}, \vb$, $\mathcal{O}$ $\rightarrow \mathcal{M}$  \tcp*{MILP model $\mathcal{M}$ for node $x$}

@@

		$6\times500$\  &   $26\%$ & $32.5\%$ & $41\%$ & $44.5\%$  & $\mathbf{65\%}$ & $92\%$ \\

		$\epsilon = 0.035$ & 34s & 119s & 18.4s & 576s & 925s &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}
	\vspace{-0.8cm}
\end{table}
